---
title: "CKME136 W17 Capstone Project"
author: "Leung Ming Kevin Tam ID 500209198"
date: "April 12, 2017"
output: pdf_document
---

```{r IMPORTDATA, echo=TRUE}
tax06.data<-read.csv(file="property_tax_report_csv2006.csv",header=T,sep=",")
tax07.data<-read.csv(file="property_tax_report_csv2007.csv",header=T,sep=",")
tax08.data<-read.csv(file="property_tax_report_csv2008.csv",header=T,sep=",")
tax09.data<-read.csv(file="property_tax_report_csv2009.csv",header=T,sep=",")
tax10.data<-read.csv(file="property_tax_report_csv2010.csv",header=T,sep=",")
tax11.data<-read.csv(file="property_tax_report_csv2011.csv",header=T,sep=",")
tax12.data<-read.csv(file="property_tax_report_csv2012.csv",header=T,sep=",")
tax13.data<-read.csv(file="property_tax_report_csv2013.csv",header=T,sep=",")
tax14.data<-read.csv(file="property_tax_report_csv2014.csv",header=T,sep=",")
tax15.data<-read.csv(file="property_tax_report_csv2015.csv",header=T,sep=",")
tax16.data<-read.csv(file="property_tax_report_csv2016.csv",header=T,sep=",")
tax17.data<-read.csv(file="property_tax_report.csv",header=T,sep=",")
```
check for NA on TAX_ASSESSMENT_YEAR fields
```{r CHECKNA, echo=TRUE}
sum(is.na(tax06.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax07.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax08.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax09.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax10.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax11.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax12.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax13.data$TAX_ASSESSMENT_YEAR))
```
confirm that 5566 records for tax14.data, 5478 records for tax15.data, 3734 records for tax16.data, 1024 records for tax17.data are missing values of TAX_ASSESSMENT_YEAR column
```{r NUMOFNA, echo=TRUE}
sum(is.na(tax14.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax15.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax16.data$TAX_ASSESSMENT_YEAR))
sum(is.na(tax17.data$TAX_ASSESSMENT_YEAR))
```
replacing missing values for TAX_ASSESSMENT_YEAR field
```{r REPLACENA, echo=TRUE}
tax14.data$TAX_ASSESSMENT_YEAR[is.na(tax14.data$TAX_ASSESSMENT_YEAR)] <- 2014
tax15.data$TAX_ASSESSMENT_YEAR[is.na(tax15.data$TAX_ASSESSMENT_YEAR)] <- 2015
tax16.data$TAX_ASSESSMENT_YEAR[is.na(tax16.data$TAX_ASSESSMENT_YEAR)] <- 2016
tax17.data$TAX_ASSESSMENT_YEAR[is.na(tax17.data$TAX_ASSESSMENT_YEAR)] <- 2017
```
binding all datasets
```{r COMBINEDATASETS, echo=TRUE}
tax.data <- rbind(tax06.data,tax07.data)
tax.data <- rbind(tax.data,tax08.data)
tax.data <- rbind(tax.data,tax09.data)
tax.data <- rbind(tax.data,tax10.data)
tax.data <- rbind(tax.data,tax11.data)
tax.data <- rbind(tax.data,tax12.data)
tax.data <- rbind(tax.data,tax13.data)
```
add back missing columns from the datasets of tax06.data to tax13.data to matching up with those of tax14.data to tax17.data
```{r ADDCOL, echo=TRUE}
tax.data$ZONE_NAME <- NA
tax.data$ZONE_CATEGORY <- NA
tax.data$PREVIOUS_LAND_VALUE <- NA
tax.data$PREVIOUS_IMPROVEMENT_VALUE <- NA
```
binding the rest
```{r echo=TRUE}
tax.data <- rbind(tax.data,tax14.data)
tax.data <- rbind(tax.data,tax15.data)
tax.data <- rbind(tax.data,tax16.data)
tax.data <- rbind(tax.data,tax17.data)
```
remove records with empty CURRENT_LAND_VALUE
```{r REMOVENA, echo=TRUE}
sum(is.na(tax.data$CURRENT_LAND_VALUE))
tax.data <- tax.data[!(is.na(tax.data$CURRENT_LAND_VALUE)),]
sum(is.na(tax.data$CURRENT_LAND_VALUE))
```
Appraoch 1
create simple linear regression with one independent variable, TAX_ASSESSMENT_YEAR (x) to predict the future value of CURRENT_LAND_VALUE (y)
```{r XYvec, echo=TRUE}
x <- as.vector(tax.data$TAX_ASSESSMENT_YEAR)
y <- as.vector(tax.data$CURRENT_LAND_VALUE)
```
create formula with Linear Regression for the prediction to CURRENT_LAND_VALUE and plot
```{r LM, echo=TRUE}
lm.price = lm(formula = y ~ x, data = tax.data)
lm.price
plot(x,y, xlab="Year", ylab="LANDVALUE")
abline(lm.price)
```
Predicated CURRENT_LAND_VALUE for 2018 is
```{r echo=TRUE}
-161818411 + (2018 * 80907)
```
Predicated CURRENT_LAND_VALUE for 2019 is
```{r echo=TRUE}
-161818411 + (2019 * 80907)
```
Predicated CURRENT_LAND_VALUE for 2020 is
```{r echo=TRUE}
-161818411 + (2020 * 80907)
```
can see some extreme outliers
```{r echo=TRUE}
summary(tax.data$CURRENT_LAND_VALUE)
boxplot(tax.data$CURRENT_LAND_VALUE)
```
replace outliers with NA and remove them
```{r echo=TRUE}
tax.data$CURRENT_LAND_VALUE <- ifelse(tax.data$CURRENT_LAND_VALUE %in% boxplot.stats(tax.data$CURRENT_LAND_VALUE)$out, NA, tax.data$CURRENT_LAND_VALUE)
tax.data <- tax.data[!(is.na(tax.data$CURRENT_LAND_VALUE)),]
sum(is.na(tax.data$CURRENT_LAND_VALUE))
boxplot(tax.data$CURRENT_LAND_VALUE)
```
re-create the vectors
```{r reXYvec, echo=TRUE}
x <- as.vector(tax.data$TAX_ASSESSMENT_YEAR)
y <- as.vector(tax.data$CURRENT_LAND_VALUE)
```
re-create formula with Linear Regression for the prediction to CURRENT_LAND_VALUE and plot
```{r LM-approach1, echo=TRUE}
lm.price = lm(formula = y ~ x, data = tax.data)
lm.price
plot(x,y, xlab="Year", ylab="LANDVALUE")
abline(lm.price)
```
Predicated CURRENT_LAND_VALUE for 2018, 2019 and 2020 are:
```{r echo=TRUE}
predict_years <- data.frame(x = c(2018, 2019, 2020))
predict(lm.price, predict_years)
```
Approach 1a
extract the first 3 characters of PROPERTY_POSTAL_CODE as FSA (Forward Sortation Area) and use it to group the predicted land value by neighbourhood
```{r echo=TRUE}
tax.dataFSA <- tax.data
tax.dataFSA$FSA <- substr(tax.dataFSA$PROPERTY_POSTAL_CODE, 1, 3)
```
validating FSAs and removing invalid or non-Vancouver FSAs
```{r echo=TRUE}
unique(tax.dataFSA$FSA,,)
FSAtoremove <- c("", "Z 8", "Z 9", "Z 2", "Z 4", "Z15", "Z 3", "Z 6", "Z 1", "M5W", "Z16", "Z10", "Z 5", " V5")
aFSAtmp <- unique(tax.dataFSA$FSA,,)
idxFSAtoremove = which(aFSAtmp %in% FSAtoremove)
aFSA <- aFSAtmp[-idxFSAtoremove]
```
separate the dataset by FSA and build Linear regression for each of them, then Predicated CURRENT_LAND_VALUE for 2018, 2019 and 2020 are:
```{r LM-approach1a, echo=TRUE}
for (FSA in aFSA) {
  newdf <- paste("df.", FSA, sep = "")
  assign(newdf, tax.dataFSA[tax.dataFSA$FSA == FSA,])
}

newdf <- NULL
newlm <- NULL
predict_years <- data.frame(TAX_ASSESSMENT_YEAR = c(2018, 2019, 2020))
for (i in 1:29) {
  newdf[i] <- paste("df.", aFSA[i], sep = "")
  assign(newdf[i], tax.dataFSA[tax.dataFSA$FSA == aFSA[i],])
  newlm[i] <- paste("lm.", aFSA[i], sep = "")
  assign(newlm[i], lm(formula = CURRENT_LAND_VALUE ~ TAX_ASSESSMENT_YEAR, data = get(newdf[i])))
  print(newlm[i])
  print(get(newlm[i]))
  newlm_tmp <- get(newlm[i])
  newpredict <- predict(newlm_tmp, predict_years)
  print(newpredict)
}

```
Approach 2
use only last 4 year datasets for multiple regression as first 8 year datasets do not even have PREVIOUS_LAND_VALUE demographics
```{r echo=TRUE}
tax.data1 <- rbind(tax14.data,tax15.data)
tax.data1 <- rbind(tax.data1,tax16.data)
tax.data1 <- rbind(tax.data1,tax17.data)
```
remove records with empty CURRENT_LAND_VALUE and PREVIOUS_LAND_VALUE
```{r REMOVENA2, echo=TRUE}
sum(is.na(tax.data1$CURRENT_LAND_VALUE))
sum(is.na(tax.data1$PREVIOUS_LAND_VALUE))
tax.data1 <- tax.data1[!(is.na(tax.data1$CURRENT_LAND_VALUE)),]
tax.data1 <- tax.data1[!(is.na(tax.data1$PREVIOUS_LAND_VALUE)),]
sum(is.na(tax.data1$CURRENT_LAND_VALUE))
sum(is.na(tax.data1$PREVIOUS_LAND_VALUE))
```
building another regression with multiple indepedent variable 
```{r LM-approach2, echo=TRUE}
lm.price1 = lm(formula = CURRENT_LAND_VALUE ~ TAX_ASSESSMENT_YEAR + PREVIOUS_LAND_VALUE, data = tax.data1)
lm.price1
```
extract mean and median of CURRENT_LAND_VALUE from 2016 dataset for fitting the following regression prediction
```{r echo=TRUE}
meanCLV2016 <- mean(tax.data$CURRENT_LAND_VALUE[tax.data$TAX_ASSESSMENT_YEAR == 2016])
medianCLV2016 <- median(tax.data$CURRENT_LAND_VALUE[tax.data$TAX_ASSESSMENT_YEAR == 2016])
```
Predicated CURRENT_LAND_VALUE for 2018, 2019 and 2020 and for the mean/median of 2016's CURRENT_LAND_VALUE are:
```{r echo=TRUE}
predict_years_CLV1 <- data.frame(TAX_ASSESSMENT_YEAR = c(2018, 2019, 2020), PREVIOUS_LAND_VALUE = meanCLV2016)
predict(lm.price1, predict_years_CLV1)
predict_years_CLV2 <- data.frame(TAX_ASSESSMENT_YEAR = c(2018, 2019, 2020), PREVIOUS_LAND_VALUE = medianCLV2016)
predict(lm.price1, predict_years_CLV2)
```

Approach 2a
built another sets of multiple regressions with grouping of FSA, and Predicated CURRENT_LAND_VALUE for 2018, 2019 and 2020 and for the mean/median of 2016's CURRENT_LAND_VALUE are:
```{r LM-approach2a, echo=TRUE}
tax.data1FSA <- tax.data1
tax.data1FSA$FSA <- substr(tax.data1FSA$PROPERTY_POSTAL_CODE, 1, 3)

newdf1 <- NULL
newlm1 <- NULL
for (i in 1:29) {
  newdf1[i] <- paste("df1.", aFSA[i], sep = "")
  assign(newdf1[i], tax.data1FSA[tax.data1FSA$FSA == aFSA[i],])
  newlm1[i] <- paste("lm1.", aFSA[i], sep = "")
  assign(newlm1[i], lm(formula = CURRENT_LAND_VALUE ~ TAX_ASSESSMENT_YEAR + PREVIOUS_LAND_VALUE, data = get(newdf1[i])))
  print(newlm1[i])
  print(get(newlm1[i]))
  newlm_tmp <- get(newlm[i])
  newpredict1 <- predict(newlm_tmp, predict_years_CLV1)
  print(newpredict1)
  newpredict2 <- predict(newlm_tmp, predict_years_CLV2)
  print(newpredict2)
}
```
Comparison of residuals of each regression
Apporach 1 and 2
```{r rmse+mse_simple, echo=TRUE}
tax.data <- tax.data[!(is.na(tax.data$CURRENT_LAND_VALUE)),]
price.actual = tax.data$CURRENT_LAND_VALUE
price.predicted = fitted(lm.price)
price.error = price.actual-price.predicted
price.rmse = sqrt(mean(price.error^2))
price.rmse
price.mae = mean(abs(price.error))
price.mae

price1.actual = tax.data1$CURRENT_LAND_VALUE
price1.predicted = fitted(lm.price1)
price1.error = price1.actual-price1.predicted
price1.rmse = sqrt(mean(price1.error^2))
price1.rmse
price1.mae = mean(abs(price1.error))
price1.mae
```
Comparison of residuals of each regression
Apporach 1a and 2a
```{r rmse+mse_multiple, echo=TRUE}
newactual <- NULL
newpredicted <- NULL
newerror <- NULL
newrmse <- NULL
newmae <- NULL
for (i in 1:29) {
  newactual[i] <- paste("price.actual.", aFSA[i], sep = "")
  assign(newactual[i], get(newdf[i])$CURRENT_LAND_VALUE)
  newpredicted[i] <- paste("price.predicted.", aFSA[i], sep = "")
  newlm_tmp <- get(newlm[i])
  assign(newpredicted[i], fitted(newlm_tmp))
  newerror[i] <- paste("price.error.", aFSA[i], sep = "")
  assign(newerror[i], get(newactual[i])-get(newpredicted[i]))
  newrmse[i] <- paste("price.rmse.", aFSA[i], sep = "")
  assign(newrmse[i], sqrt(mean(get(newerror[i])^2)))
  print(newrmse[i])
  print(get(newrmse[i]))
  newmae[i] <- paste("price.mae.", aFSA[i], sep = "")
  assign(newmae[i], mean(abs(get(newerror[i]))))
  print(newmae[i])
  print(get(newmae[i]))
}

newactual1 <- NULL
newpredicted1 <- NULL
newerror1 <- NULL
newrmse1 <- NULL
newmae1 <- NULL
for (i in 1:29) {
  newactual1[i] <- paste("price1.actual.", aFSA[i], sep = "")
  assign(newactual1[i], get(newdf1[i])$CURRENT_LAND_VALUE)
  newpredicted1[i] <- paste("price1.predicted.", aFSA[i], sep = "")
  newlm_tmp <- get(newlm1[i])
  assign(newpredicted1[i], fitted(newlm_tmp))
  newerror1[i] <- paste("price1.error.", aFSA[i], sep = "")
  assign(newerror1[i], get(newactual1[i])-get(newpredicted1[i]))
  newrmse1[i] <- paste("price1.rmse.", aFSA[i], sep = "")
  assign(newrmse1[i], sqrt(mean(get(newerror1[i])^2)))
  print(newrmse1[i])
  print(get(newrmse1[i]))
  newmae1[i] <- paste("price1.mae.", aFSA[i], sep = "")
  assign(newmae1[i], mean(abs(get(newerror1[i]))))
  print(newmae1[i])
  print(get(newmae1[i]))
}